# (PART) دروس العربي {-}

# نحو{#nahw}

نحو


```{r echo=FALSE}
# library(DiagrammeR)
# graph_kalimat <- create_graph(directed = F) %>% 
#   add_node(label = "إسم") %>% add_node(label = "معرفة/نكرة") %>% 
#   add_node(label = "مذكر/مؤنث") %>% add_node(label = "رقم") %>% 
#   add_edge(from = 1, to = 2) %>% add_edge(from = 1, to = 3) %>% add_edge(from = 1, to = 4) %>% 
#   add_node(label = "فعل") %>% 
#   add_node(label = "ماضي/مضارع\nامر/نهي") %>% add_node(label = "معروف/مجهول") %>% 
#   add_edge(from = 5, to = 6) %>% add_edge(from = 5, to = 7) %>% 
#   add_node(label = "حرف") %>% 
#   add_node(label = "كلمة/مفرد") %>% 
#   add_edge(from = 1, to = 9) %>% add_edge(from = 5, to = 9) %>% add_edge(from = 8, to = 9) %>% 
#   add_node(label = "تام") %>% add_node(label = "ناقص") %>% add_node(label = "جملة/مركب") %>% 
#   add_edge(from = 10, to = 12) %>% add_edge(from = 11, to = 12) %>% 
#   add_node(label = "موضوع") %>% add_node(label = "مهمل") %>% add_node(label = "لفظ") %>% 
#   add_edge(from = 12, to = 13) %>% add_edge(from = 9, to = 13) %>% 
#   add_edge(from = 13, to = 15) %>% add_edge(from = 14, to = 15) %>% 
#   set_node_attrs(node_attr = "font", values = "Amiri Quran")
```

Arabic words can be classified starting from the vocalization of sounds (Figure \@ref(fig:nahw-chart)). A sound is called *لفظ*. A meaningless sound, such as the vocalizations of a baby, are called *مهمل*. A sound with meaning is called *موضوغ*. This can be further classified into *جملة* (sentence) or *مركب* (compound word), which can subsequently be classified into *مفيد* (complete) or *غير مفيد* (incomplete). موضوع is also classified into *كلمة*, or word.  
إفسام الإسم:

- **جامد** An إسم that does not come from a 3-letter root word. *lit.* solid. Examples: شمس, فرس
- **مصدر** The root word. 
- **مشتق** A derived word

```{r nahw-chart, echo=FALSE, fig.cap="Chart of Arabic word taxonomy"}
library(DiagrammeR)
grViz("digraph flowchart {
      # node definitions with substituted label text
      node [fontname = Arial, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      tab7 [label = '@@7']
      tab8 [label = '@@8']
      tab9 [label = '@@9']
      tab10 [label = '@@10']
      tab11 [label = '@@11']
      tab12 [label = '@@12']
      tab13 [label = '@@13']
      tab14 [label = '@@14']
      tab15 [label = '@@15']
      tab16 [label = '@@16']
      tab17 [label = '@@17']
      tab18 [label = '@@18']
      tab19 [label = '@@19']

      # edge definitions with the node IDs
      tab1 -> tab2;
      tab1 -> tab3 -> tab4 -> tab5;
      tab4 -> tab6;
      tab3 -> tab7 -> tab15 -> tab14;
      tab15 -> tab13;
      tab15 -> tab12;
      tab7 -> tab11 -> tab10;
      tab11 -> tab9;
      tab7 -> tab8;
      tab6 -> tab16;
      tab6 -> tab17;
      tab15 -> tab18;
      tab8 -> tab19;
      }

      [1]: 'لفظ'
      [2]: 'مهمل'
      [3]: 'موضوع'
      [4]: 'جملة/مركب'
      [5]: 'غير مفيد'
      [6]: 'مفيد'
      [7]: 'كلمة/مفرد'
      [8]: 'حرف'
      [9]: 'معروف/مجهول'
      [10]: 'ماضي/مضارع/امر/نهي'
      [11]: 'فعل'
      [12]: 'رقم'
      [13]: 'مذكر/مؤنث'
      [14]: 'معرفة/نكرة'
      [15]: 'إسم'
      [16]: 'خبرية'
      [17]: 'إنشائية'
      [18]: 'جامد/مصدر/مشتق'
      [19]: 'عامل/غير عامل'
      ")
```

***

This page last compiled on `r Sys.time()`.
